---
title: "ESM 244 HW 01"
author: "Laura Gray"
date: "January 27, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##0. Load Packages and Data

```{r load, include = FALSE}

library(tidyverse)
library(ggplot2)
library(dplyr)
library(factoextra)

agencies <- read_csv("agencies.csv")
View(agencies)

launches <- read_csv("launches.csv")
View(launches)

usda <- read_csv("usda_nutrients.csv")
View(usda)

```

Launches:

- JD: Julian  Date of launch
- type: type of launch vehicle
- category: success (O) or failure (F)
- state_code: this will probably be useful in linking to "agencies" df

Abbreviated List of State Codes:
- US = United States
- SU = Soviet Union
- RU = Russia
- J = Japan
- F = France
- CN = China
- IN = India

##1. Explore "launches" through Inital Graphs

```{r}

ggplot(launches, aes(x=launch_year, y=state_code)) +
  geom_point(aes(fill=category)) +
  geom_jitter() +
  theme_classic()
# Conclusions: wow, the jitterplot reveals in a very overwhelming way that "US", "SU", "RU", "F", "CN" all have many launches per year. There are interesting things I could do with that information, now... It also looks like the Soviet Union just turned into Russia as far as state codes are concerned, so I may want to merge those designations.  I should also bind data from the agencies df to the launches df so that actual names of programs appear instead of just their state codes. 
# Questions: why is color fill not working for F/O distinction?

ggplot(launches, aes(x = state_code)) +
  geom_histogram(aes(fill = category), stat = "count") +
  theme_classic()
#facet_wrap(~launch_year)
# Conclusion: Now I can start to see something about successess/failures! Nice to know that failures is a small-looking proportion. Can also see those top few programs represented here, again, in terms of gross numbers of launches.
# Questions: still haven't seen anything of the temporal aspect. Would be interesting to see how launches changed over time, in total numbers but also in terms of successes/failures. Did the success of launches improve over time?
```

##2. Explore "agencies" through Inital Graphs

```{r}


ggplot(agencies, aes(x=state_code)) +
  geom_histogram(aes(fill=agency_type), stat="count")
# Conclusion: not a super interesting graph; can also get this information from the launches df


```

##3. Wrangle this to be better.

```{r}

#exploratory graph copied from above
ggplot(top5states, aes(x=launch_year, y=state_code)) +
  geom_point() +
  geom_jitter() +
  theme_classic()

top6states <- filter(launches, state_code == "US"| state_code == "F"| state_code == "RU"| state_code == "SU"| state_code == "CN"| state_code == "IN")

#change state code names, and merge SU and RU
top5states <- top6states %>%  
  mutate(state_code = ifelse(state_code == "RU", "Russia/USSR", as.character(state_code))) %>% 
  mutate(state_code = ifelse(state_code == "CN", "China", as.character(state_code))) %>% 
  mutate(state_code = ifelse(state_code == "SU", "Russia/USSR", as.character(state_code))) %>% 
  mutate(state_code = ifelse(state_code == "US", "United States", as.character(state_code))) %>% 
  mutate(state_code = ifelse(state_code == "IN", "India", as.character(state_code))) %>% 
  mutate(state_code = ifelse(state_code == "F", "France", as.character(state_code)))

ggplot(top5states, aes(x = state_code)) +
  geom_histogram(aes(fill = agency_type), stat = "count") +
  theme_classic()

#more wrangling... i want summed launches for each year to graph anything interesting
tally01 <- top5states %>% 
  group_by(launch_year, state_code, category) %>% 
  tally() 
View(tally01)

tally02success <- tally01 %>% 
  filter(category == "O")
View(tally02success)

ggplot(tally02success, aes(x = launch_year, y = n, color = state_code)) + 
  #geom_point() +
  scale_color_manual(values = c("red","skyblue", "orange","black","blue")) + 
  geom_line(alpha=0.7, size=1) +
  theme_classic()


####### FINISHED IN NEW MARKDOWN DOCUMENT = HW02b #######

```

##4. PCA for raw fruit and vegetable nutrients 

A. Wrangle the data to only include food groups 'Vegetables and Vegetable Products' and 'Fruits and Fruit Juices'

B. Only retain observations for which the short description (ShortDescrip) contains the string 'RAW'. Hint: see function str_detect() in the stringr package (it's in the tidyverse) 

C.  Perform principle components analysis based on variables from 'Protein_g' through 'Zinc_mg'. If you haven't added or removed columns from the original data frame, those are columns 9:30.  

D. Create a nice biplot showing both individual points (unlabeled), variables (labeled, repelled), with the two food groups (Fruits, Vegetables) differentiated by color. You'll probably want to use factoextra::fviz_pca_biplot
 with customization. You do not need to include a figure caption. For more examples, HERE is a good starting point. 

E. In text below your graph, describe several general trends that you notice from the biplot re: relationships between variables (just pick 2 - 3 pairs, and describe what the bipot tells you about the correlation) and between the food groups. 

```{r}

# filter data for vegetables and fruits (food group), incl. 'raw' (short description), select rows 9:30
fruitveg01 <- usda %>% 
  filter(FoodGroup == "Vegetables and Vegetable Products" | FoodGroup == "Fruits and Fruit Juices") %>% 
  filter(str_detect(ShortDescrip, "RAW")) 
  #select(2,9:30)
View(fruitveg01)

# perform pca
fruitveg01_pca <- prcomp(fruitveg01[9:30], scale = TRUE)
summary(fruitveg01_pca)

# create a biplot
fviz_pca_biplot(fruitveg01_pca, label = c("var"), repel = TRUE)

#couldn't solve this one: habillage = 2

#if(nrow(ind)!=length(habillage)) 
#stop("The number of active individuals is different ",
#             "from the length of the factor habillage. Please, remove the supplementary ",
#             "individuals in the variable habillage.")
```




















